---
title: "Writing_function"
output: html_document
date: "2025-10-28"
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load key packages:

```{r}
library(tidyverse)
library(readxl)
library(rvest)
```


## Start small

z-score:

```{r}
x_vec = rnorm(20, mean = 10, sd = 3.5)

(x_vec - mean(x_vec)) / sd(x_vec)
```


Write a function to compute z scores:

```{r}
z_score = function(x){
  
  if(!is.numeric(x)){
    stop("The input x should be numeric")
  }
  
  if(length(x) < 5){
    stop("Only compute z scores when the input has 5 or more numbers")
  }
  
  z = (x - mean(x)) / sd(x)
  
  z
}

```


Try the function:

```{r}
z_score(x = x_vec)

num_vec = rnorm(123, mean = 14, sd = 0.4)

z_score(x = num_vec)
```


Let's break our function:

```{r, error = TRUE}
z_score(3)
z_score("hello, ")
```


## let's compute:

let's compute the mean and sd of a numeric vector

```{r}
mean_and_sd = function(x){
  
    if(!is.numeric(x)){
    stop("The input x should be numeric")
  }
  
  if(length(x) < 5){
    stop("Only compute mean and sd when the input has 5 or more numbers")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)
  
  
  tibble(
    mean = mean_x,
    sd = sd_x
  )
}
```

## Make up data:

Let's *simulate* some data

```{r}
sim_df = 
  tibble(
    x = rnorm(n = 30, mean = 3, sd = 2)
  )

sim_df |> 
  summarize(
    mu_hat = mean(x),
    sigma_hat = sd(x)
  )
```

Write a function to do simulations

The inputs are:

* `n_subj` is the number of objects
* `mu` is the true mean
* `sigma` is the true sd

Function simulates data from normal and computes sample mean and sd

```{r}
sim_mean_sd = function(n_subj = 30, mu = 3, sigma = 2){
  
  sim_df = 
  tibble(
    x = rnorm(n = n_subj, mean = mu, sd = sigma)
  )

  sim_df |> 
  summarize(
    mu_hat = mean(x),
    sigma_hat = sd(x)
  )
  
}
```

Let's run this function.

```{r}
sim_mean_sd(mu = 48)
```


Import the lord of rings data:

```{r}
fellowship_ring = 
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") |> 
  mutate(movie = "Fellowship of the Ring")

Two_towers = 
  read_excel("data/LotR_Words.xlsx", range = "F3:H6") |> 
  mutate(movie = "Two Towers")

return_of_the_king = 
  read_excel("data/LotR_Words.xlsx", range = "J3:L6") |> 
  mutate(movie = "Return of the King")

lotr_df = 
  bind_rows(fellowship_ring, Two_towers, return_of_the_king)
```

Turn to a function:

```{r}
lotr_import_df = function(cell_range, movie_title){
  
  df = read_excel("data/LotR_Words.xlsx", range = cell_range) |> 
  mutate(movie = movie_title)
  
  df  
}

fellowship = lotr_import_df(cell_range = "B3:D6", movie_title = "Fellowship")
two_towers = lotr_import_df(cell_range = "F3:H6", movie_title = "Two towers")
return = lotr_import_df(cell_range = "J3:L6", movie_title = "Return")

bind_rows(fellowship, two_towers, return)
```


Another example

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj_year = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

data_marj_month = 
  nsduh_html |> 
  html_table() |> 
  nth(2) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

data_marj_first = 
  nsduh_html |> 
  html_table() |> 
  nth(3) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

```


Write an import function

```{r}
nsduh_import = function(html, table_num){
  
  data = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  data
}

example = nsduh_import(nsduh_html, table_num = 1)
```


